Review Paper Draft

### 2020-07-01 main.tex
3.6.4 edits

 Our initial hypothesis was that structures of $\bgamma$ that can be well-approximated by a linear combination of the observed genetic data would be well-approximated and corrected for by PC-Lasso, whereas LMM-Lasso may be more able to correct for subpopulation-specific effects that are nonlinear in nature and as such do not lend themselves to such an approximation. To that end we investigated two forms of environmental effect magnitudes: (1) exponential, and (2) dichotomous. \anna{more details?} In addition to the magnitudes of effects, we also considered whether individuals in genetically similar subpopulations would experience similar environmental effects. For example, a geographic environmental exposure such as the air quality described in our motivating example is likely to be similar among genetically related individuals due to \anna{spatial population drift, cline... (???)}. However, more behavior-based environmental exposures may be less likely to be shared by genetically similar subpopulations. Consider cigarette smoking rates in France, England, and China. While French and English populations are more genetically similar than French and Chinese populations \anna{need citaiton}, the environmental exposure of cigarette smoking is more similar between French and Chinese populations due to cultural norms and practices \anna{need citaiton}.

 To evaluate the performance of penalized LMM methods in both scenarios, we consider exponential and dichotomous effect magnitudes with are (1) \anna{homogeneous/cis (?) - same direction}, or (2) \anna{heterogeneous/trans (?)} in nature for a total of four distinct $\bgamma$ structures. We define \anna{homogenous/cis} effects as those that are similar among genetically similar subpopulations, and \anna{heterogeneous/trans} effects as those that are not. Figure \ref{fig:gamma_structures} displays a visual summary of these coefficient structures. \anna{is this still unclear? Would a PCA picture help?}
\begin{figure}[H]
    \centering
    \includegraphics[scale = 0.9]{figures/gamma_structure.png}
    \caption{$\bgamma$ structures}
    \label{fig:gamma_structures}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[scale = 1.1]{figures/beta_mse.png}
    \caption{MSE for different data types in the presence and absence of environmental confounding. \anna{$\gamma$ structure is half-and-half-decreasing-heterogeneous, subpop size is large}}
    \label{fig:mse}
\end{figure}

@@

with the y-axis displaying the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures. The dashed horizontal line at 0 indicates no difference in the performance of PC-Lasso and LMM-Lasso, while differences greater than 0 correspond to smaller MSE for LMM-Lasso, and thus superior estimation accuracy. In the presence of a small number of large subpopulations, there is some overlap in the estimation accuracy of PC-Lasso and LMM-Lasso, as evidenced by the corresponding box straddling the 0 line. However, the majority of the box falls below 0, indicating the superior performance of PC-Lasso in many of the simulations. In the presence of many small subpopulations, however, LMM-Lasso is clearly superior to PC-Lasso in estimation accuracy. This is consistent with \citep{hoffman2013correcting}, who showed that principal component adjustment methods represent a low dimensional approximation to linear mixed models. As such, it is reasonable that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform well in the small-number-of-large-subpopulations setting, but fall short in the presence of many small subpopulations. 
\begin{figure}[H]
    \centering
    % \includegraphics[scale = 0.9]{figures/mse_diff_subpops}
     \includegraphics[scale = 0.9]{figures/mse_diff_subpops.png}
    \caption{Relative performance of LMM-Lasso, PC-Lasso in terms of estimation accuracy:  \anna{structureX is independent subpops, $\bgamma$ structure is dichotomous and heterogeneous, $\eta= 0.5, \xi$ is varying}}
    \label{fig:big_vs_small}
\end{figure}

@@

\begin{figure}[H]
    \centering
    % \includegraphics[scale = 0.9]{figures/fig2b.png}
    \includegraphics[scale = 0.9]{figures/mse_diff_hetero.png}
    \caption{Relative performance of LMM-Lasso, PC-Lasso for different subpopulation and $\bgamma$ structures: \anna{structureX is independent subpops, $\eta= 0.5, \xi = 0.8$}}
    \label{fig:big_small_gamma}
\end{figure}

@@

\anna{Fix factor label greek letters - R update broke some things}
\begin{figure}[H]
    \centering
    \includegraphics[scale = 1]{figures/eta_beta_hat.png}
     \caption{Caption goes here \anna{$\xi = 0.5$, structureX = independent subpops, structureGamma is dichotomous heterogeneous, large blocks}}
    \label{fig:eta_beta_mse}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[scale = 0.8]{figures/time_ratio.png}
    \caption{Caption goes here \anna{$\eta = 0.5, \xi = 0.8$, structureX = independent subpops, structureGamma is dichotomous heterogeneous, large blocks}}
    \label{fig:time}
\end{figure}

>>>>

 Our initial hypothesis was that structures of $\bgamma$ that can be well-approximated by a linear combination of the observed genetic data would be well-approximated and corrected for by PC-Lasso, whereas LMM-Lasso may be more able to correct for subpopulation-specific effects that are nonlinear in nature and as such do not lend themselves to such an approximation. To that end we investigated two forms of environmental effect magnitudes: (1) exponential, and (2) dichotomous. \anna{more details?} In addition to the magnitudes of effects, we also considered whether individuals in genetically similar subpopulations would experience similar environmental effects. For example, a geographic environmental exposure such as the air quality described in our motivating example is likely to be similar among genetically related individuals due to concordance between genetic clusters and geography. However, behavior-based environmental exposures may be less likely to be shared by subpopulations which are genetically and geographically similar. \anna{another citation?} As an example, \citep{ng2014smoking} found considerable within-region variation in smoking rates in Africa, Asia, and South America.

 \anna{in organic chemistry cis means 'on the same side' and trans 'on the other side' so I thought cis/trans might be more descriptive for gamma structure than homo/heterogeneous. These are also used to describe gender identity, which is obviously not how I'm using it, so I don't want it to be confusing.}
 To evaluate the performance of penalized LMM methods in both scenarios, we consider exponential and dichotomous effect magnitudes with are (1) \anna{homogeneous/cis}, or (2) \anna{heterogeneous/trans} in nature for a total of four distinct $\bgamma$ structures. We define \anna{homogenous/cis} effects as those that are similar among genetically similar subpopulations, and \anna{heterogeneous/trans} effects as those that are not. Figure \ref{fig:gamma_structures} displays a visual summary of these coefficient structures. \anna{is this still unclear? Would a PCA picture help?}
\begin{figure}[H]
    \centering
    \includegraphics[scale = 0.9]{figures/gamma_structure.png}
    \caption{$\bgamma$ structures}
    \label{fig:gamma_structures}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[scale = 1.1]{figures/beta_mse.png}
    \caption{Estimation accuracy of different methods for different data types in the presence and absence of environmental confounding with coarse subpopulation structure and dichotomous-\anna{heterogeneous/trans} environmental effects.}
    \label{fig:mse}
\end{figure}

@@

with the y-axis displaying the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures. The dashed horizontal line at 0 indicates no difference in the performance of PC-Lasso and LMM-Lasso, while differences greater than 0 correspond to smaller MSE for LMM-Lasso, and thus superior estimation accuracy. In the presence of a small number of large subpopulations, there is some overlap in the estimation accuracy of PC-Lasso and LMM-Lasso, as evidenced by the corresponding box straddling the 0 line. However, the majority of the box falls below 0, indicating the superior performance of PC-Lasso in many of the simulations. In the presence of many small subpopulations, however, LMM-Lasso is clearly superior to PC-Lasso in estimation accuracy. This is consistent with \citep{hoffman2013correcting}, who showed that principal component adjustment methods represent a low dimensional approximation to linear mixed models. As such, it is reasonable that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform well in the small-number-of-large-subpopulations setting, but fall short in the presence of many small subpopulations. 
\begin{figure}[H]
    \centering
    % \includegraphics[scale = 0.9]{figures/mse_diff_subpops}
     \includegraphics[scale = 0.9]{figures/mse_diff_subpops.png}
    \caption{Relative performance of LMM-Lasso, PC-Lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with independent subpopulation data, $\eta = 0.5$, and dichotomous-\anna{heterogeneous/trans} environmental effects.}
    \label{fig:big_vs_small}
\end{figure}

@@

\begin{figure}[H]
    \centering
    % \includegraphics[scale = 0.9]{figures/fig2b.png}
    \includegraphics[scale = 0.9]{figures/mse_diff_hetero.png}
    \caption{Relative performance of LMM-Lasso, PC-Lasso for fine and coarse subpopulation structures and varying environmental effect structures with independent subpopulation data, $\eta = 0.5$, and $\xi = 0.8$}
    \label{fig:big_small_gamma}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[scale = 1]{figures/eta_beta_hat.png}
     \caption{$\bbeta$ and $\eta$ estimation accuracy of LMM-Lasso-Rakitsch and LMM-Lasso-ggmix for varying $\eta$ levels and sample sizes with independent subpopulation data, coarse subpopulation structure, $\xi = 0.8$, and dichotomous-\anna{heterogeneous/trans} environmental effect structure.}
    \label{fig:eta_beta_mse}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[scale = 0.8]{figures/time_ratio.png}
    \caption{Computing time of LMM-Lasso-Rakitsch compared with that of LMM-Lasso-ggmix for varying sample sizes with independent subpopulation structure, coarse subpopulation structure, $\eta = 0.5$, $\xi = 0.8$, and dichotomous-\anna{heterogeneous/trans} environmental effect structure.}
    \label{fig:time}
\end{figure}
